---
title: "The Effects of Moisture Stress on Organic Soil"
author: "Sarah Gao"
date: "2023-03-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Context

This Rmarkdown file is used to bring in and clean each data type (i.e. total CN, qPCR DNA, aqueous N, and CO2 respiration). It keeps all non-error data, such as blanks and negative controls, during cleaning, without any standardization, calibration, or normalization.


```{r packages}
# Load relevant packages

library("dplyr")
library("readr")
library("tidyr")
library("fs")

```

```{r functions}
# Source necessary functions
source("code/functions/ea_functions/clean_ea_data.R")
source("code/functions/qpcr_functions/clean_qpcr_data.R")
source("code/functions/n_functions/n_clean_n_data.R")

```

## Total CN EA Data

The following chunk reads and cleans Total CN data from the ThermoFisher EA runs, specifically percentage CN and ratio C:N in separate data files. Due to the finickiness of the EA machine, there were several runs that were labelled as SRM or sample numbers that had a value of 0. These are omitted, but all blanks (i.e. system blanks with no capsules, blanks that were just a tin capsule, and blanks that were just a silver capsule) are retained.s 

```{r ea-data}
# This code chunk reads in the raw EA total CN data and cleans it

# Compile list of file paths of percentage EA data
files_percent <- dir_ls(path = "data/raw_data/EA_CN/2022/",
                        recurse = 1,
                        regex =
                          "\\w+_Run\\d_(repeat_)*(\\d{2}_)*percent.(xls|XLS)")
# Compile list of file paths of C:N ratio data
files_ratio <- dir_ls(path = "data/raw_data/EA_CN/2022/",
                      recurse = 1,
                      regex = "\\w+_Run\\d_(repeat_)*(\\d{2}_)*ratio.(xls|XLS)")

# Clean both types of data
# Note that the function filters out any sample or SRM runs that had a value of 0. Blanks are retained.
cn_percent_clean <- clean_ea_data(files_percent)
cn_ratio_clean <- clean_ea_data(files_ratio)
```

## qPCR DNA Data

Here, raw qPCR fungal and bacterial DNA data is read and cleaned. All blank runs are retained as well, which may result in NAs for `rep_no` and even `cq` values. There are some non-blank sample wells where Cq values were NA due to unknown reasons. In fungal samples, these included the water-only samples (32, 42, 78, 97, 119) but there were also some samples that had a stray replicate that was NA when other replicates amplified.
```{r qpcr-dna}
# Get clean fungal data
clean_fungal <- clean_qpcr_data("data/raw_data/qPCR/", "fungal")

# Get clean bacterial data
clean_bacterial <- clean_qpcr_data("data/raw_data/qPCR/", "bacterial")
```

Here, dried soil weight measurements are read and cleaned. These weights were derived from frozen soil weights measured for DNA extraction. For each sample, approximately the same amount of frozen soil was later weighed out and then oven dried. These dried weights are reported here.

The dried weights data also includes qubit concentration data taken after DNA extraction to determine the DNA concentrations of the resulting extractions.

Samples 32, 42, 78, 97, and 119 were water-only controls (i.e. only water in jars throughout the experiment) and thus has NA values for soil weights. In the `extraction_soil_wt_mg` column for these water-only jars, reported weights are the amount of water added to the DNA extraction.

For the `bacteria-control` sample, a sample taken from bacteria cultured on agar from a soil sample was used for DNA extraction to serve as a positive control. The `extraction_soil_wt_mg` value is the weight of the bacteria sample (including any agar media) used for DNA extraction.
```{r qpcr-dried-wts}
# Read in dried soil weights data
dried_wts <- read_csv("data/raw_data/qPCR/2022_DNA.csv") %>%
  select(-"...5")
```

## Inorganic N Data

This code chunk reads in and cleans data from inorganic N measurements. These include both the extractable N measurements as well as the leachate N measurements. Measured species were ammonia (NH3) and combined nitrite-nitrate (NO2-NO3).

Reported measurements are already calibrated to NO2, NO3, and ammonia stock standards. These standards are retained in this dataset and are shown as `NO2E6`, `NO3E6`, and `NH3E` respectively. Blank values (i.e. technical blanks that underwent the same extraction / leachate process with the same solutions but without any added soil) are also shown.

The column in the dataset labelled `OD` stands for the raw optical density measurement. Replicate number (`rep_no`) reflects the technical replicate number used during the SmartChem analysis (i.e. these replicates were created during SmartChem analysis and use the same sample). For blanks, these replicate numbers may be duplicated as blanks were taken during each round of sample collection. The labeling for blanks in `sample_no_full`, which is omitted in the final cleaned dataset, is coded to the batch of solution and is pretty haphazard. For example, `Blank1AL02` reflects the results of the second technical replicate of a leachate blank made with batch 1A of CaCl2. `BlankE102` is the second technical replicate of an extract blank made with batch 1 of KCl. There are multiples of each `BlankE1` replicate due to the fact that I made one big 5L batch of KCl that lasted for the entire experiment whereas I had to make multiple new batches of the CaCl2 solution.

Some samples have measurements of `-9999` for unknown reasons, which have been retained in this dataset.
```{r inorg-n}
# Get file list
file_list <- paste0("data/raw_data/SmartChem_N_extractions/2022_samples/",
                    list.files(
                      "data/raw_data/SmartChem_N_extractions/2022_samples/",
                        pattern = "*.Csv"))

# Read in and clean all csv files in the folder and create a compiled dataframe
n_data_clean <- clean_n_data(file_list)
```

## CO2 Respiration Data

This code chunk
```{r co2}

```
